## Loading Libraries

```{r message=FALSE, warning=FALSE}
library(readr)
library(readxl)
library(janitor)
library(tidyverse)
library(broom)
library(lme4)
library(lmerTest)
library(plotly)
library(knitr)
library(cowplot)
```

## Reading in Data and Cleaning
```{r message=FALSE, warning=FALSE}
# get economic level data per dissemination area
economic_level_per_da <- read_excel("data/raw/Data sheets & inventory.xlsx", sheet='Overall Analysis') |>
  select(`Economic level (k)...21`:`DA...22`) |>
  drop_na() |>
  rename(economic_level_k = `Economic level (k)...21`,
         dissemination_area = `DA...22`)

plant_shrub_full_data <- read_delim("data/raw/plant_shrub_data.csv", delim=';') |>
  clean_names() |>
  fill(house, .direction = "down")

bird_data <- read_excel("data/raw/BIRDS_spreadsheet2.xlsx", sheet="Long_clean") |>
  clean_names()

# get native percentage
garden_data <- read_excel("data/raw/GARDENS_spreadsheet.xlsx", sheet='Native_vs_not_GRAPHING') |>
  select(House:Economic_level) |> # select necessary columns
  drop_na() |>  # remove empty rows
  clean_names()  # clean column names

native_percent <- garden_data |> # get native percentage per house
  select(house, native_percent_total)

# get main data
stats_data <- read_excel("data/raw/Data sheets & inventory.xlsx", sheet='Everything_statistics') |>
  select(House:Tree_percentage) |>  # select necessary columns
  drop_na() |>  # remove empty rows
  clean_names() |>  # clean column names
  mutate(dissemination_area = as.factor(str_sub(house, 1, 2))) |> # add dissemination area
  left_join(native_percent, by = join_by(house)) # add native percentage
```
## OSFL, BARS, NOFL, PIWO, CBCH, CEDW, VATH, PAWR, RBNU

### Time of day detection
```{r}
bird_subset_data <- bird_data |>
  filter(species %in% c("OSFL", "BARS", "NOFL", "PIWO", "CBCH", "CEDW", "VATH", "PAWR", "RBNU"))

ggplot(bird_subset_data, aes(x = hour)) +
  geom_histogram(binwidth = 1, fill="skyblue", colour="white") +
  facet_wrap(~species, scales = "free_y", axes = "all_x") +
  scale_x_continuous(breaks = 4:12) +
  theme_minimal() +
  labs(
    title = "Detection Time Distribution by Species",
    x = "Hour of Day",
    y = "Number of Detections"
  )
```

Per House:
```{r, fig.height=10}
ggplot(bird_subset_data, aes(x = hour, y = house, colour = house)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Species of Interest Detection Time Per House",
       x = "Hour of Day",
       y = "House") +
  facet_wrap(~species, scales = "free_y", axes = "all_x")
```

### Economic level
```{r, fig.width=8, fig.height=6}
bird_house_data <- bird_subset_data |>
  distinct(house, species)
  
bird_stats_data <- stats_data |>
  filter(house %in% bird_house_data$house) |>
  full_join(bird_house_data, by="house")

ggplot(bird_stats_data, aes(x=economic_level, y= species)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(colour = species), show.legend = FALSE) +
  theme_minimal() +
  labs(y = "Species",
       x = "Economic Level (k)",
       title = "Economic Level of Areas where Species of Interest Were Detected")
```

### Canopy Cover
```{r, fig.width=8, fig.height=15}
garden_canopy_plot <- ggplot(bird_stats_data, aes(x=canopy_cover_garden, y=species)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(colour = species), show.legend = FALSE) +
  theme_minimal() +
  labs(y = "Species",
       x = "Canopy Cover (m^2)",
       title = "Garden Canopy Cover of Areas where Species of Interest were Detected")

canopy_50_plot <- ggplot(bird_stats_data, aes(x=canopy_cover_radius_50m, y=species)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(colour = species), show.legend = FALSE) +
  theme_minimal() +
  labs(y = "Species",
       x = "Canopy Cover Radius (m)",
       title = "Canopy Cover Radius 50m where Species of Interest were Detected")

canopy_100_plot <- ggplot(bird_stats_data, aes(x=canopy_cover_radius_100m, y=species)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(colour = species), show.legend = FALSE) +
  theme_minimal() +
  labs(y = "Species",
       x = "Canopy Cover Radius (m)",
       title = "Canopy Cover Radius 100m where Species of Interest were Detected")

plot_grid(garden_canopy_plot, canopy_50_plot, canopy_100_plot, ncol = 1)
```

### Overall Garden Diversity
```{r, fig.width=8, fig.height=6}
ggplot(bird_stats_data, aes(x=overall_garden_diversity, y=species)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(colour = species), show.legend = FALSE) +
  theme_minimal() +
  labs(y = "Species",
       x = "Overall Garden Diversity",
       title = "Garden Diversity of Areas where Species of Interest were Detected")
```

### Native Plant Percentage
```{r, fig.width=8, fig.height=6}
ggplot(bird_stats_data, aes(x=native_percent_total, y=species)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(colour = species), show.legend = FALSE) +
  theme_minimal() +
  labs(y = "Species",
       x = "Native Vegetation (%)",
       title = "Native Vegetation Percentage of Areas where Species of Interest were Detected")
```


## GAMS

Suggested GAMs:
1. detection probability GAM with canopy cover or vegetation diversity
2. presence probability vs native plant%

→ I think we can do a logistic regression model to model the presence or absence of a species using predictors such as canopy cover, vegetation diversity, and native plant %. We can use an step wise approach or other approach to select our predictors, and the model output should tell us which predictors are actually significant.

However, we might not be able to do this for species that are either present at nearly every house or absent from nearly all houses.


Quick look at number of unique houses per species:
```{r}
n_houses_per_species <- bird_subset_data |>
  group_by(species) |>
  summarise(n_houses = n_distinct(house)) |>
  arrange(n_houses)

n_houses_per_species
```

### Logistic Regression Model: Olive-Sided Flycatcher (OSFL)
```{r}
# get data necessary for modeling
plant_shrub_full_data <- read_csv("data/processed/cleaned_plant_shrub_data.csv")

stats_garden_data <- left_join(stats_data, garden_data, by="house", suffix = c("", ".y")) |>
  select(-ends_with(".y"))

osfl_houses <- bird_stats_data |>
  filter(species == "OSFL") |>
  distinct(house)

plant_shrub_data_summarized <- plant_shrub_full_data |>
  group_by(house) |>
  summarize(
    number_fruiting_species = sum(fruiting),
    number_flowering_species = sum(flowering),
    number_grain_species = sum(grain),
  )

stats_garden_data <- left_join(stats_garden_data, plant_shrub_data_summarized, by="house")

osfl_stats_garden_data <- stats_garden_data |>
  mutate(osfl_present = house %in% osfl_houses$house)
```
```{r}
ggplot(osfl_stats_garden_data) +
  geom_point(aes(economic_level, osfl_present)) +
  geom_smooth(aes(economic_level, osfl_present),
    method = "glm", formula = y ~ x,
    method.args = c(family = binomial), se = FALSE, color = "blue"
  )
```

```{r}
osfl_model <- glmer(
  osfl_present ~
    economic_level +
    number_fruiting_species +
    number_flowering_species +
    number_grain_species +
    (1 | dissemination_area),
  data = osfl_stats_garden_data,
  family = binomial
)

summary(osfl_model)$coefficients |>
  as.data.frame() |>
  mutate(odds_estimate = exp(Estimate),
         odds_estimate_inverted = 1/odds_estimate) |>
  kable()
```

### Logistic Regression Model: Barn Swallow (BARS)
```{r}
bars_stats_garden_data <- stats_garden_data |>
  mutate(bars_present = house %in% osfl_houses$house)

bars_model <- glmer(
  bars_present ~
    economic_level +
    number_fruiting_species +
    number_flowering_species +
    number_grain_species +
    (1 | dissemination_area),
  data = bars_stats_garden_data,
  family = binomial
)

summary(bars_model)$coefficients |>
  as.data.frame() |>
  mutate(odds_estimate = exp(Estimate),
         odds_estimate_inverted = 1/odds_estimate) |>
  kable()
```