# Results

```{r message = FALSE, warning = FALSE}
# Loading libraries and data
library(readxl)
library(janitor)
library(MASS)
library(AER)
library(tidyverse)
library(broom)
library(lme4)
library(lmerTest)
library(plotly)
library(knitr)
library(dplyr)

GARDEN_FILE_PATH <- "data/raw/GARDENS_spreadsheet.xlsx"
garden_data <- read_excel(GARDEN_FILE_PATH, sheet='Everything_statistics') |>
  clean_names() # clean column names

# File path to save images of the graphs to
IMAGE_FILE_PATH <-"images/"
```

## Chapter 1

### Canopy cover in backyards vs. Economic Level

```{r}
canopy_ecomomic_lr <- lm(canopy_cover_garden_maybepercentage ~ economic_level, garden_data)
results_canopy_ecomomic_lr <- summary(canopy_ecomomic_lr)
results_canopy_ecomomic_lr
```

```{r}
canopy_ecomomic_lr_pvalue <- results_canopy_ecomomic_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
canopy_ecomomic_lr_r2 <- results_canopy_ecomomic_lr$r.squared |>
  round(4)

cat("p-value:", canopy_ecomomic_lr_pvalue, "\n")
cat("R-squared:", canopy_ecomomic_lr_r2)
```

```{r, message = FALSE}
canopy_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = canopy_cover_garden_maybepercentage)) +
  geom_smooth(method = "lm") +
  geom_point(data = garden_data[-17, ]) +
  theme_minimal() +
  labs(x = "Economic level",
       y = "Canopy cover garden (%)",
       title = "Backyard canopy cover vs Economic level")

canopy_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-14_canopy-cover-vs-economic-level.png"),
       plot = canopy_ecomomic_plot)
```

### Tree total biomass per backyard vs. Economic Level

```{r}
tree_biomass_ecomomic_lr <- lm(biomass_total_kg ~ economic_level, garden_data)
results_tree_biomass_ecomomic_lr <- summary(tree_biomass_ecomomic_lr)
results_tree_biomass_ecomomic_lr
```

```{r}
tree_biomass_ecomomic_lr_pvalue <- results_tree_biomass_ecomomic_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
tree_biomass_ecomomic_lr_r2 <- results_tree_biomass_ecomomic_lr$r.squared |>
  round(4)

cat("p-value:", tree_biomass_ecomomic_lr_pvalue, "\n")
cat("R-squared:", tree_biomass_ecomomic_lr_r2)
```

```{r, message = FALSE}
garden_biomass_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = biomass_total_kg)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic level",
       y = "Garden biomass (kg)",
       title = "Garden biomass vs Economic level")

garden_biomass_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-15_garden-biomass-vs-economic-level.png"),
       plot = garden_biomass_ecomomic_plot)
```

### Total vegetation species richness per backyard vs Economic level

```{r}
veg_richness_ecomomic_lr <- lm(overall_garden_diversity ~ economic_level, garden_data)
results_veg_richness_lr <- summary(veg_richness_ecomomic_lr)
results_veg_richness_lr
```

```{r}
veg_richness_ecomomic_lr_pvalue <- results_veg_richness_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
veg_richness_lr_r2 <- results_veg_richness_lr$r.squared |>
  round(4)

cat("p-value:", veg_richness_ecomomic_lr_pvalue, "\n")
cat("R-squared:", veg_richness_lr_r2)
```

```{r, message = FALSE}
veg_richness_plot <-
  ggplot(garden_data, aes(x = economic_level, y = overall_garden_diversity)) +
  geom_smooth(method = "lm") +
  geom_point(data = garden_data[-1, ]) +
  theme_minimal() +
  labs(x = "Economic level",
       y = "Total backyard plant species richness",
       title = "Total backyard plant species richness vs Economic level")

veg_richness_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-16_veg-richness-vs-economic-level.png"),
       plot = veg_richness_plot)
```

### Tree diversity in tems of percentage area per backyeard vs. Economic Level

```{r}
shannon_ecomomic_lr <- lm(shannon_index_per_garden_withpercentcover ~ economic_level, garden_data)
results_shannon_ecomomic_lr <- summary(shannon_ecomomic_lr)
results_shannon_ecomomic_lr
```

```{r}
shannon_ecomomic_lr_pvalue <- results_shannon_ecomomic_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
shannon_biomass_ecomomic_lr_r2 <- results_shannon_ecomomic_lr$r.squared |>
  round(4)

cat("p-value:", shannon_ecomomic_lr_pvalue, "\n")
cat("R-squared:", shannon_biomass_ecomomic_lr_r2)
```

```{r, message = FALSE}
shannon_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level,
                          y = shannon_index_per_garden_withpercentcover)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic level",
       y = "Shannon Diversity Index",
       title = "Shannon Diversity in terms of Canopy percent cover vs Economic level")

shannon_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-18_shannon-diversity-vs-economic-level.png"),
       plot = shannon_ecomomic_plot)
```

### ANOVA on how economic level influences tree species richness
```{r}
# add economic group level names to data
garden_data <- garden_data |>
  mutate(economic_category = as.factor(str_sub(house, 1, 1)),
         economic_category = as.factor(
           case_when(
             economic_category == "P" ~ "Low",
             economic_category == "M" ~ "Medium",
             economic_category == "R" ~ "High"
             )
           ),
         # specify order of economic group levels
         economic_category = fct_relevel(economic_category, c("Low", "Medium", "High")))

tree_richness_economic_anova <- aov(tree_species_richness ~ economic_category, garden_data)
results_tree_richness_economic_anova <- summary(tree_richness_economic_anova)
results_tree_richness_economic_anova
```
```{r}
tree_richness_economic_anova_df <-
  results_tree_richness_economic_anova[[1]][[1, "Df"]] |>
  round(4)

tree_richness_economic_anova_df_resid <-
  results_tree_richness_economic_anova[[1]][[2, "Df"]] |>
  round(4)

tree_richness_economic_anova_fvalue <-
  results_tree_richness_economic_anova[[1]][[1, "F value"]] |>
  round(4)

tree_richness_economic_anova_pvalue <- 
  results_tree_richness_economic_anova[[1]][[1, "Pr(>F)"]] |>
  round(4)


cat("df (degrees of freedom):", tree_richness_economic_anova_df, "\n")
cat("Residual:", tree_richness_economic_anova_df_resid, "\n")
cat("F:", tree_richness_economic_anova_fvalue, "\n")
cat("p-value:", tree_richness_economic_anova_pvalue)
```

```{r}
tree_richness_economic_plot <-
  ggplot(garden_data, aes(x = economic_category, y = tree_species_richness)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Economic level",
       y = "Tree species richness",
       title = "Tree species richness across economic levels")

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-19_tree-richness-vs-economic-level.png"),
       plot = tree_richness_economic_plot)
```

## Chapter 2

### Bird species richness in each backyard vs % canopy cover in back yard
removed outlier in analysis
```{r}
# (omit outlier)
bird_canopy_lr <- lm(bird_diversity ~ canopy_cover_garden_maybepercentage, garden_data[-13, ])
results_bird_canopy_lr <- summary(bird_canopy_lr)
results_bird_canopy_lr
```

```{r}
bird_canopy_lr_pvalue <-
  results_bird_canopy_lr$coefficients["canopy_cover_garden_maybepercentage", "Pr(>|t|)"] |>
  round(4)
bird_canopy_lr_r2 <- results_bird_canopy_lr$r.squared |>
  round(4)

cat("p-value:", bird_canopy_lr_pvalue, "\n")
cat("R-squared:", bird_canopy_lr_r2)
```

```{r, message = FALSE}
bird_canopy_plot <-
  ggplot(garden_data[-13, ], aes(x = canopy_cover_garden_maybepercentage,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Canopy cover backyard (%)",
       y = "Bird species richness",
       title = "Bird species richness vs Backyard canopy cover")

bird_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-21_bird_vs_canopy.png"),
       plot = bird_canopy_plot)
```

### Bird species richness vs vegetation species richness
kept outlier in analysis
```{r}
# (omit outlier)
bird_veg_lr <- lm(bird_diversity ~ overall_garden_diversity, garden_data)
results_bird_veg_lr <- summary(bird_veg_lr)
results_bird_veg_lr
```

```{r}
bird_veg_lr_pvalue <-
  results_bird_veg_lr$coefficients["overall_garden_diversity", "Pr(>|t|)"] |>
  round(4)
bird_veg_lr_r2 <- results_bird_veg_lr$r.squared

cat("p-value:", bird_veg_lr_pvalue, "\n")
cat("R-squared:", bird_veg_lr_r2)
```

```{r, message = FALSE}
bird_veg_plot <-
  ggplot(garden_data[-13, ], aes(x = overall_garden_diversity,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Total plant species richness in backyard",
       y = "Bird species richness",
       title = "Bird species richness vs Backyard plant diversity")

bird_veg_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-21_bird_vs_veg_richness.png"),
       plot = bird_veg_plot)
```