# Analysis

```{r message = FALSE, warning = FALSE}
# Loading libraries and data
library(readxl)
library(janitor)
library(MASS)
library(tidyverse)
library(broom)
library(purrr)
library(cowplot)
library(car)
library(AER)

GARDEN_FILE_PATH <- "data/raw/GARDENS_spreadsheet.xlsx"
garden_data <- read_excel(GARDEN_FILE_PATH, sheet='Everything_statistics') |>
  clean_names() # clean column names

coniferous_data <- read_excel(GARDEN_FILE_PATH, sheet='Tree_per_garden') |>
  clean_names() |>
  filter(coniferous_not == "con") |>
  group_by(id) |>
  summarise(coniferous_canopy = sum(crown_area))

garden_data <- left_join(garden_data, coniferous_data, by = c("house" = "id")) |>
  mutate(coniferous_canopy = replace_na(coniferous_canopy, 0))

BIRD_FILE_PATH <- "data/raw/BIRDS_spreadsheet2.xlsx"
bird_data <- read_excel(BIRD_FILE_PATH, sheet="Long_clean") |>
  clean_names()

# File path to save images of the graphs to
IMAGE_FILE_PATH <-"images/"
```

## Chapter 1

To investigate how economic level is associated with the structure and composition of vegetation in private urban forests, linear models were fitted with economic level as the independent variable and the following response variables per backyard: canopy cover (with an outlier with very high canopy cover omitted), tree total biomass, and Shannon diversity index based on canopy percent cover.

In addition, a negative binomial regression model was fitted to assess whether total vegetation species richness was associated with economic level. As the data were overdispersed, a negative binomial model was used instead of Poisson. An outlier with very high vegetation species richness was omitted from the analysis.

A one-way ANOVA was used to test whether tree species richness differed across economic groups. Normality and homogeneity of variances were verified using the Shapiro-Wilk and Levene's tests, respectively.

### Backyard Canopy Cover vs. Economic Level

```{r}
no_p22_data <- garden_data |>
  filter(house != "P22")

canopy_ecomomic_lr <- lm(canopy_cover_garden_maybepercentage ~ economic_level, no_p22_data)
results_canopy_ecomomic_lr <- summary(canopy_ecomomic_lr)
results_canopy_ecomomic_lr
```
p-value: 0.000126

R-squared: 0.4379

```{r message = FALSE, warning = FALSE}
canopy_ecomomic_plot <-
  ggplot(no_p22_data, aes(x = economic_level, y = canopy_cover_garden_maybepercentage)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Canopy Cover Garden (%)",
       title = "Backyard Canopy Cover vs. Economic Level")

canopy_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-14_canopy-cover-vs-economic-level.png"),
       plot = canopy_ecomomic_plot)
```
Example figure caption: Relationship between backyard canopy cover (%) and economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

### Backyard Tree Biomass vs. Economic Level

```{r}
tree_biomass_ecomomic_lr <- lm(biomass_total_kg ~ economic_level, garden_data)
results_tree_biomass_ecomomic_lr <- summary(tree_biomass_ecomomic_lr)
results_tree_biomass_ecomomic_lr
```

p-value: 0.0459

R-squared: 0.1348

```{r, message = FALSE}
garden_biomass_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = biomass_total_kg)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Backyard Tree Biomass (kg)",
       title = "Backyard Tree Biomass vs. Economic Level")

garden_biomass_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-15_tree-biomass-vs-economic-level.png"),
       plot = garden_biomass_ecomomic_plot)
```
Example figure caption: Relationship between bacyard tree biomass (kg) and economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

### Total Backyard Plant Species Richness vs. Economic Level

Fit Poisson model since vegetation species richness is a count variable and checking for overdispersion. A Poisson model assumes equidispersion, where the variance of the response variable (vegetation species richness) is equal to its mean. In the case that its variance is greater than its mean, a negative binomial model should be used instead.
```{r}
no_m11_data <- garden_data |>
  filter(house != "M11")

veg_richness_ecomomic_poisson <-
  glm(overall_garden_diversity ~ economic_level,
      family = poisson,
      data = no_m11_data)
dispersiontest(veg_richness_ecomomic_poisson)
```
Since the dispersion test yields p-value = 0.001 < 0.05, there is overdispersion in the data and we'll use a negative binomial model.

```{r}
veg_richness_ecomomic_nb <-
  glm.nb(overall_garden_diversity ~ economic_level,
      data = no_m11_data)
results_veg_richness_nb <- tidy(veg_richness_ecomomic_nb, exponentiate = TRUE)
results_veg_richness_nb
```
p-value = 0.5549

```{r, message = FALSE}
veg_richness_plot <-
  ggplot(no_m11_data, aes(x = economic_level, y = overall_garden_diversity)) +
  geom_smooth(method = "glm.nb") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Total Backyard Plant Species Richness",
       title = "Total Backyard Plant Species Richness vs. Economic Level")

veg_richness_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-16_veg-richness-vs-economic-level.png"),
       plot = veg_richness_plot)
```
Example figure caption: Relationship between plant species richness and economic level. The line represents the negative binomial model fit with 95% confidence interval (shaded grey).

### Shannon Diversity in Terms of Canopy Percent Cover vs. Economic Level

```{r}
shannon_ecomomic_lr <- lm(shannon_index_per_garden_withpercentcover ~ economic_level, garden_data)
results_shannon_ecomomic_lr <- summary(shannon_ecomomic_lr)
results_shannon_ecomomic_lr
```
p-value: 0.00485

R-squared: 0.2505

```{r, message = FALSE}
shannon_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level,
                          y = shannon_index_per_garden_withpercentcover)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Shannon Diversity Index",
       title = "Shannon Diversity in Terms of Canopy Percent Cover vs. Economic Level")

shannon_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-18_shannon-diversity-vs-economic-level.png"),
       plot = shannon_ecomomic_plot)
```
Example figure caption: Relationship between Shannon Diversity Index in terms of canopy percent cover and economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

### ANOVA on Tree Species Richness Across Economic Groups
```{r}
# add economic group level names to data
garden_data <- garden_data |>
  mutate(economic_category = as.factor(str_sub(house, 1, 1)),
         economic_category = as.factor(
           case_when(
             economic_category == "P" ~ "Low",
             economic_category == "M" ~ "Medium",
             economic_category == "R" ~ "High"
             )
           ),
         # specify order of economic group levels
         economic_category = fct_relevel(economic_category, c("Low", "Medium", "High")))
```

ANOVA:
```{r}
tree_richness_economic_anova <- aov(tree_species_richness ~ economic_category, garden_data)
results_tree_richness_economic_anova <- summary(tree_richness_economic_anova)
results_tree_richness_economic_anova
```
df (degrees of freedom): 2

Residual: 27

F: 4.409

p-value: 0.022


**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(tree_richness_economic_anova$residuals)
```

```{r}
leveneTest(tree_species_richness ~ economic_category, garden_data)
```
Neither assumption is violated (p-values > 0.05).

Plot:
```{r fig.height = 5, fig.width = 5}
tree_richness_economic_plot <-
  ggplot(garden_data, aes(x = economic_category, y = tree_species_richness)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  labs(x = "Economic Group",
       y = "Tree Species Richness",
       title = "Tree Species Richness Across Economic Groups")

tree_richness_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-19_tree-richness-vs-economic-level.png"),
       plot = tree_richness_economic_plot,
       height = 5,
       width = 5)
```
Example figure caption: Tree species richness observed per economic group. Boxplots represent the mean (red diamond), median (bold horizontal line), interquartile range (boxes), and 1.5 interquartile range (whiskers).

## Chapter 2

To investigate how bird species richness is associated with vegetation composition and structure in private urban forests, Poisson regression models were fitted with bird species richness as the response variable and the following predictors per backyard: canopy cover, plant species richness, and native plant proportion. Equidispersion was verified using a dispersion test, confirming that Poisson models were appropriate. Three points were excluded from the analysis due to low bird detections attributed to recording issues.

A bird species accumulation curve was constructed by randomly ordering houses and cumulatively recording the number of new species detected with each additional house sampled. This process was repeated 100 times with different random orderings, and the mean cumulative species richness along with the 95% confidence interval across permutations were calculated.

Vegetation and bird variables, both raw and standardized per backyard area, were tested using linear models with economic level as the independent variable.

### Bird Species Richness vs. Backyard Canopy Cover
```{r}
no_p21_m13_m11 <- garden_data |>
  filter(!house %in% c("P21", "M13", "M11"))

bird_canopy_poisson <-
  glm(bird_diversity ~ canopy_cover_garden_maybepercentage,
      family = poisson,
      data = no_p21_m13_m11)
dispersiontest(bird_canopy_poisson)
```
p-value = 0.2004 > 0.05, there is no significant overdispersion.

```{r}
results_bird_canopy_poisson <- tidy(bird_canopy_poisson, exponentiate = TRUE)
results_bird_canopy_poisson
```
p-value: 0.7883

```{r, message = FALSE, warning = FALSE}
bird_canopy_plot <-
  ggplot(no_p21_m13_m11, aes(x = canopy_cover_garden_maybepercentage,
                          y = bird_diversity)) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  geom_point() +
  theme_minimal() +
  labs(x = "Canopy Cover Backyard (%)",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Canopy Cover")

bird_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-21_bird-vs-canopy.png"),
       plot = bird_canopy_plot)
```
Example figure caption: Relationship between bird species richness and backyard canopy cover (%). The line represents the Poisson model fit with 95% confidence interval (shaded grey).

### Bird Species Richness vs. Backyard Plant Species Richness
```{r}
bird_veg_poisson <-
  glm(bird_diversity ~ overall_garden_diversity,
      family = "poisson",
      data = no_p21_m13_m11)
dispersiontest(bird_veg_poisson)
```
p-value = 0.1946 > 0.05, there is no significant overdispersion.

```{r}
results_bird_veg_poisson <- tidy(bird_veg_poisson, exponentiate = TRUE)
results_bird_veg_poisson
```
p-value = 0.7407

```{r, message = FALSE}
bird_veg_plot <-
  ggplot(no_p21_m13_m11, aes(x = overall_garden_diversity,
                          y = bird_diversity)) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  geom_point() +
  theme_minimal() +
  labs(x = "Total Plant Species Richness in Backyard",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Plant Species Richness")

bird_veg_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-22_bird-vs-veg-richness.png"),
       plot = bird_veg_plot)
```
Example figure caption: Relationship between bird species richness and plant species richness. The line represents the Poisson model fit with 95% confidence interval (shaded grey).

### Bird Species Richness vs. Native Plant Proportion
```{r message = FALSE}
# add native percentage to garden data
native_percent <- read_excel(GARDEN_FILE_PATH, sheet='Native_vs_not_GRAPHING') |>
  select(House, Native_percent_total) |>
  clean_names() # clean column names

no_p21_m13_m11 <- no_p21_m13_m11 |>
  left_join(native_percent, by = join_by(house))
```

```{r}
bird_native_poisson <-
  glm(bird_diversity ~ native_percent_total,
      family = poisson,
      data = no_p21_m13_m11)
dispersiontest(bird_native_poisson)
```

p-value = 0.2123 > 0.05, there is no significant overdispersion.

```{r}
results_bird_native_poisson <- tidy(bird_native_poisson, exponentiate = TRUE)
results_bird_native_poisson
```

p-value = 0.2599


```{r, message = FALSE}
bird_native_plot <-
  ggplot(no_p21_m13_m11, aes(x = native_percent_total,
                          y = bird_diversity)) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  geom_point() +
  theme_minimal() +
  labs(x = "Native % Plants Total",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Native Plant Proportion")

bird_native_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-23_bird-vs-native.png"),
       plot = bird_native_plot)
```
Example figure caption: Relationship between bird species richness and native proportion of plants (%). The line represents the Poisson model fit with 95% confidence interval (shaded grey).

### Accumulation curve accross houses
```{r}
set.seed(123) # set seed for reproducible permutations

# presence/absence matrix where rows = houses and columns = species
presence_matrix <- bird_data |>
  count(house, species) |>
  pivot_wider(names_from = species, values_from = n, values_fill = 0) |>
  column_to_rownames("house") |>
  mutate(across(everything(), ~ as.integer(. > 0)))

# species accumulation curve
n_permutations <- 100
houses <- rownames(presence_matrix)
n_houses <- length(houses)
n_species <- ncol(presence_matrix)

curves <- matrix(0, nrow = n_permutations, ncol = n_houses)

for (i in seq_len(n_permutations)) {
  perm <- sample(houses)
  seen <- rep(FALSE, n_species)

  for (j in seq_along(perm)) {
    seen <- seen | as.logical(presence_matrix[perm[j], ])
    curves[i, j] <- sum(seen)
  }
}

# summarise across permutations
results <- tibble(
  x          = seq_len(n_houses),
  mean_curve = colMeans(curves),
  lower      = apply(curves, 2, quantile, probs = 0.025),
  upper      = apply(curves, 2, quantile, probs = 0.975)
)

# plot
accumulation_plot <- ggplot(results, aes(x = x)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey60", alpha = 0.4) +
  geom_line(aes(y = mean_curve), color = "#3366FF", linewidth = 1) +
  labs(
    title = "Species Accumulation Curve Across Houses",
    x     = "Number of Houses Sampled",
    y     = "Cumulative Species Detected"
  ) +
  theme_minimal()

accumulation_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-24_accumulation-plot.png"),
       plot = accumulation_plot)
```
Example figure caption: Bird species accumulation curve based on 100 random permutations of houses. The line represents the mean cumulative species richness and with the 95% confidence interval (shaded grey).

### Panel comparing vegetation and birds variables vs economic level and the same over area of backyard
```{r fig.height = 12, fig.width = 6, warning = FALSE, message = FALSE}
garden_data <- garden_data |>
  mutate(biomass_per_m2 = biomass_total_kg/garden_size_m_2,
         plants_per_m2 = overall_garden_diversity/garden_size_m_2,
         trees_per_m2 = tree_species_richness/garden_size_m_2,
         canopy_per_m2 = canopy_cover_garden_maybepercentage/garden_size_m_2,
         shannon_index_per_garden_withpercentcover_per_m2 = shannon_index_per_garden_withpercentcover/garden_size_m_2,
         birds_per_m2 = bird_diversity/garden_size_m_2
         )

y_vars <- c("biomass_total_kg", "biomass_per_m2",
            "overall_garden_diversity", "plants_per_m2",
            "tree_species_richness", "trees_per_m2",
            "canopy_cover_garden_maybepercentage", "canopy_per_m2",
            "shannon_index_per_garden_withpercentcover", "shannon_index_per_garden_withpercentcover_per_m2",
            "bird_diversity", "birds_per_m2")
titles <- c("Biomass", "Biomass per area",
            "Plant richness", "Plant richness per area",
            "Tree richness", "Tree richness per area",
            "Canopy cover", "Canopy cover per area",
            "Shannon Diversity", "Shannon Diversity per area",
            "Bird richness", "Bird richness per area")

plots <- map2(y_vars, titles, \(var, title) {
  # compute p-value for the slope of economic_level
  p <- tidy(lm(garden_data[[var]] ~ garden_data$economic_level))$p.value[2]
  p_label <- paste0("p = ", signif(p, 3))
  
  ggplot(garden_data, aes(x = economic_level, y = .data[[var]])) +
    geom_smooth(method = "lm") +
    geom_point() +
    theme_minimal() +
    annotate("text", x = -Inf, y = Inf, label = p_label,
             hjust = -0.1, vjust = 1.5, size = 4) +
    labs(y = NULL, x = NULL, title = title)
})

panel_plot <- plot_grid(plotlist = plots, ncol = 2)
panel_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-25_panel-plot.png"),
       plot = panel_plot,
       height = 12,
       width = 6)
```
Example figure caption: Relationship between biomass, plant species richness, canopy cover, and bird species richness (raw and area standardized per backyard area) against economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

## Other

To test whether tree height and average tree height per house differed across economic groups, normality and homogeneity of variances were assessed using the Shapiro-Wilk and Levene's tests, respectively. As normality was violated, a Kruskal-Wallis test was used as a non-parametric alternative. An outlier was removed from the analysis and plots.

### Tree height per economic group
```{r}
tree_height <- read_excel(GARDEN_FILE_PATH, sheet='Tree_per_garden') |>
  select(ID, `Ht (m)`) |>
  clean_names() |>
  left_join(garden_data[, c("house", "economic_category", "economic_level")], by = c("id" = "house")) |>
  filter(id != "P22") # remove outlier
```

ANOVA:
```{r}
tree_height_economic_anova <- aov(ht_m ~ economic_category, tree_height)
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(tree_height_economic_anova$residuals)
```

```{r}
leveneTest(ht_m ~ economic_category, tree_height)
```
The assumption of normally distributed residuals is violated (p-value < 0.05) while the assumption of variance is not (p-value > 0.05). Therefore, a Kruskal-Wallis test is more appropriate to compare the **median** of each group instead of the **mean**. The medians better reflects the central tendency in the case of outliers and skewedness. Additionally, the Kruskal-Wallis test does not make assumptions about the data distribution or variablility.

Kruskal-Wallis Test:
```{r}
kruskal.test(ht_m ~ economic_category, data = tree_height)
```
There is a statistically significant difference in median tree height across economic categories (p = 0.022 < 0.05).

```{r fig.height = 6, fig.width = 5}
tree_height_economic_plot <-
  ggplot(tree_height, aes(x = economic_category, y = ht_m)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Group",
       y = "Tree Height",
       title = "Tree Height Across Economic Groups")

tree_height_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-1_tree-height-vs-economic-level.png"),
      plot = tree_height_economic_plot,
      height = 6,
      width = 6)
```
Example figure caption: Tree height observed per economic group. Boxplots represent the mean (red diamond), median (bold horizontal line), interquartile range (boxes), and 1.5 interquartile range (whiskers).

#### Linear Regression of Tree Height vs. Economic Level
```{r}
tree_height_lr <- lm(ht_m ~ economic_level, data = tree_height)
results_tree_height_lr <- summary(tree_height_lr)
results_tree_height_lr
```
p-value: 0.0107

R-squared: 0.05938


I do not suggest using this model since there does not seem to be linearity in the coefficients, the residuals vs. fitted points does not show a cloud. Also, the Q-Q Residuals does not show normality of the error term since the points do not follow the 45 degree line.
```{r}
plot(tree_height_lr, 1)
plot(tree_height_lr, 2)
```

```{r message = FALSE, warning = FALSE}
tree_height_economic_box_plot <-
  ggplot(tree_height, aes(x = as.factor(economic_level), y = ht_m)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Tree Height",
       title = "Tree Height Across Economic Levels")

tree_height_economic_box_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-1_boxplot-tree-height-vs-economic-level.png"),
       plot = tree_height_economic_box_plot)
```
Example figure caption: Tree height observed per economic level. Boxplots represent the mean (red diamond), median (bold horizontal line), interquartile range (boxes), and 1.5 interquartile range (whiskers).

NOTE: I don't really recommend using a box plot for this, since there are a differing amount of observations per economic level, and the box plot hides individual observations.


```{r}
tree_height_ecomomic_lr_plot <-
  ggplot(tree_height, aes(x = economic_level, y = ht_m)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Tree Height",
       title = "Tree Height vs. Economic Level")

tree_height_ecomomic_lr_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-1_line-tree-height-vs-economic-level.png"),
       plot = tree_height_ecomomic_lr_plot)
```
Example figure caption: Relationship between tree height (m) and economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

### Average tree height per house per economic group
```{r}
average_tree_height_per_house <- tree_height |>
  group_by(id, economic_category, economic_level) |>
  summarize(average_tree_height = mean(ht_m))

avg_tree_height_economic_anova <-
  aov(average_tree_height ~ economic_category, average_tree_height_per_house)
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(avg_tree_height_economic_anova$residuals)
```

```{r}
leveneTest(average_tree_height ~ economic_category, data = average_tree_height_per_house)
```
Again, the assumption of normally distributed residuals is violated (p-value < 0.05) while the assumption of variance is not (p-value > 0.05). Therefore, a Kruskal-Wallis test again to compare the **median** of each group instead of the **mean**.

Kruskal-Wallis Test:
```{r}
kruskal.test(average_tree_height ~ economic_category, data = average_tree_height_per_house)
```
The median of average tree height per house did not differ significantly across economic categories (p-value = 0.064 > 0.05).


```{r fig.height = 6, fig.width = 5}
avg_tree_height_economic_plot <-
  ggplot(average_tree_height_per_house, aes(x = economic_category, y = average_tree_height)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Group",
       y = "Average Tree Height per House",
       title = "Average Tree Height per House Across Economic Groups")

avg_tree_height_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-2_average-tree-height-per-house-vs-economic-level.png"),
      plot = avg_tree_height_economic_plot,
      height = 6,
      width = 6)
```

Example figure caption: Average tree height per house observed per economic group. Boxplots represent the mean (red diamond), median (bold horizontal line), interquartile range (boxes), and 1.5 interquartile range (whiskers).

#### Linear Regression of Average Tree Height vs. Economic Level
```{r}
avg_tree_height_lr <- lm(average_tree_height ~ economic_level, data = average_tree_height_per_house)
results_avg_tree_height_lr <- summary(avg_tree_height_lr)
results_avg_tree_height_lr
```
p-value: 0.144

R-squared: 0.09882


I do not suggest using this model since there does not seem to be linearity in the coefficients, the residuals vs. fitted points does not show a cloud. The Q-Q Residuals shows normality of the error term since the points do follow the 45 degree line, but the former point still stands.
```{r}
plot(avg_tree_height_lr, 1)
plot(avg_tree_height_lr, 2)
```

```{r message = FALSE, warning = FALSE}
avg_tree_height_economic_box_plot <-
  ggplot(average_tree_height_per_house, aes(x = as.factor(economic_level), y = average_tree_height)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Average Tree Height per House",
       title = "Average Tree Height per House Across Economic Levels")

avg_tree_height_economic_box_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-2_boxplot-average-tree-height-vs-economic-level.png"),
       plot = avg_tree_height_economic_box_plot)
```
Example figure caption: Average tree height per house observed per economic level. Boxplots represent the mean (red diamond), median (bold horizontal line), interquartile range (boxes), and 1.5 interquartile range (whiskers).

NOTE: I don't really recommend using a box plot for this, since there are little amount of observations/houses per economic level, and the box plot can imply that there are more observations than there are.


```{r}
avg_tree_height_ecomomic_lr_plot <-
  ggplot(average_tree_height_per_house, aes(x = economic_level, y = average_tree_height)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Tree Height",
       title = "Tree Height vs. Economic Level")

avg_tree_height_ecomomic_lr_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-2_line-average-tree-height-vs-economic-level.png"),
       plot = avg_tree_height_ecomomic_lr_plot)
```
Example figure caption: Relationship between average tree height per house (m) and economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

### Backyard Coniferous Canopy Cover vs. Economic Level
Note: only 12 backyards have coniferous trees

```{r}
con_canopy_ecomomic_lr <- lm(coniferous_canopy ~ economic_level, no_p22_data)
results_con_canopy_ecomomic_lr <- summary(con_canopy_ecomomic_lr)
results_con_canopy_ecomomic_lr
```
p-value: 0.00739

R-squared: 0.2371

```{r message = FALSE, warning = FALSE}
con_canopy_ecomomic_plot <-
  ggplot(no_p22_data, aes(x = economic_level, y = coniferous_canopy)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Coniferous Canopy Cover",
       title = "Coniferous Canopy Cover vs. Economic Level")

con_canopy_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, 
                         "other-3_coniferous-canopy-cover-vs-economic-level.png"),
       plot = con_canopy_ecomomic_plot)
```
Example figure caption: Relationship between coniferous canopy cover and economic level. The line represents the linear model fit with 95% confidence interval (shaded grey).

### Bird Species Richness vs. Backyard Coniferous Canopy Cover
Note: only 12 backyards have coniferous trees.

```{r}
bird_con_canopy_poisson <-
  glm(bird_diversity ~ coniferous_canopy,
      family = poisson,
      data = no_p21_m13_m11)
dispersiontest(bird_con_canopy_poisson)
```
p-value = 0.1941 > 0.05, there is no significant overdispersion.

```{r}
results_bird_con_canopy_poisson <- tidy(bird_con_canopy_poisson, exponentiate = TRUE)
results_bird_con_canopy_poisson
```
p-value: 0.5946

```{r, message = FALSE, warning = FALSE}
bird_con_canopy_plot <-
  ggplot(no_p21_m13_m11, aes(x = coniferous_canopy,
                          y = bird_diversity)) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  geom_point() +
  theme_minimal() +
  labs(x = "Coniferous Canopy Cover",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Coniferous Canopy Cover")

bird_con_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-4_bird-vs-coniferous-canopy.png"),
       plot = bird_con_canopy_plot)
```
Example figure caption: Relationship between bird species richness and coniferous canopy cover. The line represents the Poisson model fit with 95% confidence interval (shaded grey).

### Bird Species Richness vs. 50m Canopy Cover
```{r}
bird_50m_canopy_poisson <-
  glm(bird_diversity ~ canopy_cover_radius_50m,
      family = poisson,
      data = no_p21_m13_m11)
dispersiontest(bird_50m_canopy_poisson)
```

p-value = 0.1886 > 0.05, there is no significant overdispersion.

```{r}
results_bird_50m_canopy_poisson <- tidy(bird_50m_canopy_poisson, exponentiate = TRUE)
results_bird_50m_canopy_poisson
```
p-value: 0.2737

```{r, message = FALSE, warning = FALSE}
bird_50m_canopy_plot <-
  ggplot(no_p21_m13_m11, aes(x = canopy_cover_radius_50m,
                          y = bird_diversity)) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  geom_point() +
  theme_minimal() +
  labs(x = "50m Canopy Cover",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. 50m Canopy Cover")

bird_50m_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-5_bird-vs-50m-canopy.png"),
       plot = bird_50m_canopy_plot)
```
Example figure caption: Relationship between bird species richness and 50m canopy cover. The line represents the Poisson model fit with 95% confidence interval (shaded grey).

### Bird Species Richness vs. 100m Canopy Cover
```{r}
bird_100m_canopy_poisson <-
  glm(bird_diversity ~ canopy_cover_radius_100m,
      family = poisson,
      data = no_p21_m13_m11)
dispersiontest(bird_100m_canopy_poisson)
```

p-value = 0.1823 > 0.05, there is no significant overdispersion.

```{r}
results_bird_100m_canopy_poisson <- tidy(bird_100m_canopy_poisson, exponentiate = TRUE)
results_bird_100m_canopy_poisson
```
p-value: 0.1872

```{r, message = FALSE, warning = FALSE}
bird_100m_canopy_plot <-
  ggplot(no_p21_m13_m11, aes(x = canopy_cover_radius_100m,
                          y = bird_diversity)) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  geom_point() +
  theme_minimal() +
  labs(x = "100m Canopy Cover",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. 100m Canopy Cover")

bird_100m_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-6_bird-vs-100m-canopy.png"),
       plot = bird_100m_canopy_plot)
```
Example figure caption: Relationship between bird species richness and 100m canopy cover. The line represents the Poisson model fit with 95% confidence interval (shaded grey).
