# Analysis

```{r message = FALSE, warning = FALSE}
# Loading libraries and data
library(readxl)
library(janitor)
library(tidyverse)
library(broom)
library(purrr)
library(cowplot)
library(car)
library(AER)
library(MASS)

GARDEN_FILE_PATH <- "data/raw/GARDENS_spreadsheet.xlsx"
garden_data <- read_excel(GARDEN_FILE_PATH, sheet='Everything_statistics') |>
  clean_names() # clean column names

BIRD_FILE_PATH <- "data/raw/BIRDS_spreadsheet2.xlsx"
bird_data <- read_excel(BIRD_FILE_PATH, sheet="Long_clean") |>
  clean_names()

# File path to save images of the graphs to
IMAGE_FILE_PATH <-"images/"
```

## Chapter 1

To investigate how economic level is associated with the structure and composition of vegetation in private urban forests, linear models were fitted with economic level as the independent variable and the following response variables per backyard: canopy cover, tree total biomass, and Shannon diversity index based on canopy percent cover.

In addition, a negative binomial regression model was fitted to assess whether total vegetation species richness was associated with economic level. As the data were overdispersed, a negative binomial model was used instead of Poisson.

A one-way ANOVA was used to test whether tree species richness differed across economic groups. Normality and homogeneity of variances were verified using the Shapiro-Wilk and Levene's tests, respectively.

### Canopy cover in backyards vs. Economic Level

```{r}
canopy_ecomomic_lr <- lm(canopy_cover_garden_maybepercentage ~ economic_level, garden_data)
results_canopy_ecomomic_lr <- summary(canopy_ecomomic_lr)
results_canopy_ecomomic_lr
```
p-value: 0.0129

R-squared: 0.2078

```{r message = FALSE, warning = FALSE}
no_p22_data <- garden_data |>
  filter(house != "P22")

canopy_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = canopy_cover_garden_maybepercentage)) +
  geom_smooth(method = "lm") +
  geom_point(data = no_p22_data) +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Canopy Cover Garden (%)",
       title = "Backyard Canopy Cover vs. Economic Level")

canopy_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-14_canopy-cover-vs-economic-level.png"),
       plot = canopy_ecomomic_plot)
```

### Tree total biomass per backyard vs. Economic Level

```{r}
tree_biomass_ecomomic_lr <- lm(biomass_total_kg ~ economic_level, garden_data)
results_tree_biomass_ecomomic_lr <- summary(tree_biomass_ecomomic_lr)
results_tree_biomass_ecomomic_lr
```

p-value: 0.0459

R-squared: 0.1348


ASK ABOUT TREE VS GARDEN BIOMASS
```{r, message = FALSE}
garden_biomass_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = biomass_total_kg)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Garden Biomass (kg)",
       title = "Garden Biomass vs. Economic Level")

garden_biomass_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-15_garden-biomass-vs-economic-level.png"),
       plot = garden_biomass_ecomomic_plot)
```

### Total vegetation species richness per backyard vs Economic level

Fit Poisson model since vegetation species richness is a count variable and checking for overdispersion. A Poisson model assumes equidispersion, where the variance of the response variable (vegetation species richness) is equal to its mean. In the case that its variance is greater than its mean, a negative binomial model should be used instead.
```{r}
veg_richness_ecomomic_poisson <-
  glm(overall_garden_diversity ~ economic_level,
      family = poisson,
      data = garden_data)
dispersiontest(veg_richness_ecomomic_poisson)
```
Since the dispersion test yields p-value = 0.02 < 0.05, there is overdispersion in the data and we'll use a negative binomial model.

```{r}
veg_richness_ecomomic_nb <-
  glm.nb(overall_garden_diversity ~ economic_level,
      data = garden_data)
results_veg_richness_nb <- tidy(veg_richness_ecomomic_nb, exponentiate = TRUE)
results_veg_richness_nb
```
p-value = 0.908

```{r, message = FALSE}
no_m11_data <- garden_data |>
  filter(house != "m11")
  
veg_richness_plot <-
  ggplot(garden_data, aes(x = economic_level, y = overall_garden_diversity)) +
  geom_smooth(method = "glm.nb") +
  geom_point(data = no_m11_data) +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Total Backyard Plant Species Richness",
       title = "Total Backyard Plant Species Richness vs. Economic Level")

veg_richness_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-16_veg-richness-vs-economic-level.png"),
       plot = veg_richness_plot)
```

### Tree diversity in tems of percentage area per backyard vs. Economic Level

```{r}
shannon_ecomomic_lr <- lm(shannon_index_per_garden_withpercentcover ~ economic_level, garden_data)
results_shannon_ecomomic_lr <- summary(shannon_ecomomic_lr)
results_shannon_ecomomic_lr
```
p-value: 0.0048

R-squared: 0.2505

```{r, message = FALSE}
shannon_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level,
                          y = shannon_index_per_garden_withpercentcover)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Shannon Diversity Index",
       title = "Shannon Diversity in Terms of Canopy Percent Cover vs. Economic Level")

shannon_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-18_shannon-diversity-vs-economic-level.png"),
       plot = shannon_ecomomic_plot)
```

### ANOVA on how economic level influences tree species richness
```{r}
# add economic group level names to data
garden_data <- garden_data |>
  mutate(economic_category = as.factor(str_sub(house, 1, 1)),
         economic_category = as.factor(
           case_when(
             economic_category == "P" ~ "Low",
             economic_category == "M" ~ "Medium",
             economic_category == "R" ~ "High"
             )
           ),
         # specify order of economic group levels
         economic_category = fct_relevel(economic_category, c("Low", "Medium", "High")))
```

ANOVA:
```{r}
tree_richness_economic_anova <- aov(tree_species_richness ~ economic_category, garden_data)
results_tree_richness_economic_anova <- summary(tree_richness_economic_anova)
results_tree_richness_economic_anova
```
df (degrees of freedom): 2

Residual: 27

F: 4.409

p-value: 0.022


**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(tree_richness_economic_anova$residuals)
```

```{r}
leveneTest(tree_species_richness ~ economic_category, garden_data)
```
Neither assumption is violated (p-values > 0.05).

Plot:
```{r fig.height = 5, fig.width = 5}
tree_richness_economic_plot <-
  ggplot(garden_data, aes(x = economic_category, y = tree_species_richness)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  labs(x = "Economic Group",
       y = "Tree Species Richness",
       title = "Tree Species Richness Across Economic Groups")

tree_richness_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-19_tree-richness-vs-economic-level.png"),
       plot = tree_richness_economic_plot,
       height = 5,
       width = 5)
```

## Chapter 2

To investigate how bird species richness is associated with vegetation composition and structure in private urban forests, Poisson regression models were fitted with bird species richness as the response variable and the following predictors per backyard: canopy cover, plant species richness, and native plant proportion.

A bird species accumulation curve was constructed by randomly ordering houses and cumulatively recording the number of new species detected with each additional house sampled. This process was repeated 100 times with different random orderings, and the mean cumulative species richness along with the 95% confidence interval across permutations were calculated.

Vegetation and bird variables, both raw and standardized per backyard area, were tested using linear models with economic level as the independent variable.

### Bird species richness in each backyard vs % canopy cover in back yard
```{r}
bird_canopy_lr <- lm(bird_diversity ~ canopy_cover_garden_maybepercentage, garden_data[-13, ])
results_bird_canopy_lr <- summary(bird_canopy_lr)
results_bird_canopy_lr
```
p-value: 0.4372

R-squared: 0.0234

```{r, message = FALSE}
no_p12_data <- garden_data |>
  filter(house != "p12")
  
bird_canopy_plot <-
  ggplot(garden_data, aes(x = canopy_cover_garden_maybepercentage,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point(data = no_p12_data) +
  theme_minimal() +
  labs(x = "Canopy Cover Backyard (%)",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Canopy Cover")

bird_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-21_bird_vs_canopy.png"),
       plot = bird_canopy_plot)
```

### Bird species richness vs vegetation species richness
```{r}
bird_veg_lr <- lm(bird_diversity ~ overall_garden_diversity, garden_data)
results_bird_veg_lr <- summary(bird_veg_lr)
results_bird_veg_lr
```
p-value: 0.9692

R-squared: 0.000054295

```{r, message = FALSE}
bird_veg_plot <-
  ggplot(garden_data, aes(x = overall_garden_diversity,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point(data = no_p12_data) +
  theme_minimal() +
  labs(x = "Total Plant Species Richness in Backyard",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Plant Species Richness")

bird_veg_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-22_bird_vs_veg_richness.png"),
       plot = bird_veg_plot)
```

### Bird species richness vs native plant species as a ratio of all species in backyard
```{r message = FALSE}
# add native percentage to garden data
native_percent <- read_excel(GARDEN_FILE_PATH, sheet='Native_vs_not_GRAPHING') |>
  select(House, Native_percent_total) |>
  clean_names() # clean column names

garden_data <- garden_data |>
  left_join(native_percent, by = join_by(house))
```
```{r}
bird_native_lr <- lm(bird_diversity ~ native_percent_total, garden_data[-13, ])
results_bird_native_lr <- summary(bird_native_lr)
results_bird_native_lr
```
p-value: 0.918

R-squared: 0.0003998



Limit y-axis to 0-30?

```{r, message = FALSE}
bird_native_plot <-
  ggplot(garden_data, aes(x = native_percent_total,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point(data = no_p12_data) +
  theme_minimal() +
  labs(x = "Native % Plants Total",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Native Plant Proportion")

bird_native_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-23_bird_vs_native.png"),
       plot = bird_native_plot)
```

### Accumulation curve accross houses
```{r}
set.seed(123) # set seed for reproducible permutations

# presence/absence matrix where rows = houses and columns = species
presence_matrix <- bird_data |>
  count(house, species) |>
  pivot_wider(names_from = species, values_from = n, values_fill = 0) |>
  column_to_rownames("house") |>
  mutate(across(everything(), ~ as.integer(. > 0)))

# species accumulation curve
n_permutations <- 100
houses <- rownames(presence_matrix)
n_houses <- length(houses)
n_species <- ncol(presence_matrix)

curves <- matrix(0, nrow = n_permutations, ncol = n_houses)

for (i in seq_len(n_permutations)) {
  perm <- sample(houses)
  seen <- rep(FALSE, n_species)

  for (j in seq_along(perm)) {
    seen <- seen | as.logical(presence_matrix[perm[j], ])
    curves[i, j] <- sum(seen)
  }
}

# summarise across permutations
results <- tibble(
  x          = seq_len(n_houses),
  mean_curve = colMeans(curves),
  lower      = apply(curves, 2, quantile, probs = 0.025),
  upper      = apply(curves, 2, quantile, probs = 0.975)
)

# plot
accumulation_plot <- ggplot(results, aes(x = x)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey60", alpha = 0.4) +
  geom_line(aes(y = mean_curve), color = "#3366FF", linewidth = 1) +
  labs(
    title = "Species Accumulation Curve Across Houses",
    x     = "Number of Houses Sampled",
    y     = "Cumulative Species Detected"
  ) +
  theme_minimal()

accumulation_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-24_accumulation_plot.png"),
       plot = accumulation_plot)
```

### Panel comparing vegetation and birds variables vs economic level and the same over area of backyard
```{r fig.height = 12, fig.width = 6}
garden_data <- garden_data |>
  mutate(biomass_per_m2 = biomass_total_kg/garden_size_m_2,
         plants_per_m2 = overall_garden_diversity/garden_size_m_2,
         trees_per_m2 = tree_species_richness/garden_size_m_2,
         canopy_per_m2 = canopy_cover_garden_maybepercentage/garden_size_m_2,
         birds_per_m2 = bird_diversity/garden_size_m_2,
         )

y_vars <- c("biomass_total_kg", "biomass_per_m2",
            "overall_garden_diversity", "plants_per_m2",
            "tree_species_richness", "trees_per_m2",
            "canopy_cover_garden_maybepercentage", "canopy_per_m2",
            "bird_diversity", "birds_per_m2")
titles <- c("Biomass", "Biomass per area",
            "Plant richness", "Plant richness per area",
            "Tree richness", "Tree richness per area",
            "Canopy cover", "Canopy cover per area",
            "Bird richness", "Bird richness per area")

plots <- map2(y_vars, titles, \(var, title) {
  # compute p-value for the slope of economic_level
  p <- tidy(lm(garden_data[[var]] ~ garden_data$economic_level))$p.value[2]
  p_label <- paste0("p = ", signif(p, 3))
  
  ggplot(garden_data, aes(x = economic_level, y = .data[[var]])) +
    geom_smooth(method = "lm") +
    geom_point() +
    theme_minimal() +
    annotate("text", x = -Inf, y = Inf, label = p_label,
             hjust = -0.1, vjust = 1.5, size = 4) +
    labs(y = NULL, x = NULL, title = title)
})

panel_plot <- plot_grid(plotlist = plots, ncol = 2)
panel_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-25_panel-plot.png"),
       plot = panel_plot,
       height = 12,
       width = 6)
```
## Other

To test whether tree height and average tree height per house differed across economic groups, normality and homogeneity of variances were assessed using the Shapiro-Wilk and Levene's tests, respectively. As normality was violated, a Kruskal-Wallis test was used as a non-parametric alternative.

### ANOVA tree height per economic level
```{r}
tree_height <- read_excel(GARDEN_FILE_PATH, sheet='Tree_per_garden') |>
  select(ID, `Ht (m)`) |>
  clean_names() |>
  left_join(garden_data[, c("house", "economic_category")], by = c("id" = "house")) |>
  filter(id != "P22") # remove outlier
```

ANOVA:
```{r}
tree_height_economic_anova <- aov(ht_m ~ economic_category, tree_height)
results_tree_height_economic_anova <- summary(tree_height_economic_anova)
results_tree_height_economic_anova
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(tree_height_economic_anova$residuals)
```

```{r}
leveneTest(ht_m ~ economic_category, tree_height)
```
The assumption of normally distributed residuals is violated (p-value < 0.05) while the assumption of variance is not (p-value > 0.05). Therefore, a Kruskal-Wallis test is more appropriate to compare the **median** of each group instead of the **mean**. The medians better reflects the central tendency in the case of outliers and skewedness. Additionally, the Kruskal-Wallis test does not make assumptions about the data distribution or variablility.

Kruskal-Wallis Test:
```{r}
kruskal.test(ht_m ~ economic_category, data = tree_height)
```
There is a statistically significant difference in median tree height across economic categories (p = 0.022 < 0.05).

```{r fig.height = 6, fig.width = 5}
tree_height_economic_plot <-
  ggplot(tree_height, aes(x = economic_category, y = ht_m)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Group",
       y = "Tree Height",
       title = "Tree Height Across Economic Groups")

tree_height_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-1_tree-height-vs-economic-level.png"),
      plot = tree_height_economic_plot,
      height = 6,
      width = 6)
```
  
### ANOVA average tree height per house per economic level
```{r}
average_tree_height_per_house <- tree_height |>
  group_by(id, economic_category) |>
  summarize(average_tree_height = mean(ht_m))

avg_tree_height_economic_anova <-
  aov(average_tree_height ~ economic_category, average_tree_height_per_house)

summary(avg_tree_height_economic_anova)
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(avg_tree_height_economic_anova$residuals)
```

```{r}
leveneTest(average_tree_height ~ economic_category, data = average_tree_height_per_house)
```
Again, the assumption of normally distributed residuals is violated (p-value < 0.05) while the assumption of variance is not (p-value > 0.05). Therefore, a Kruskal-Wallis test again to compare the **median** of each group instead of the **mean**.

Kruskal-Wallis Test:
```{r}
kruskal.test(average_tree_height ~ economic_category, data = average_tree_height_per_house)
```
The median of average tree height per house did not differ significantly across economic categories (p-value = 0.064 > 0.05).


```{r fig.height = 6, fig.width = 5}
avg_tree_height_economic_plot <-
  ggplot(average_tree_height_per_house, aes(x = economic_category, y = average_tree_height)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Group",
       y = "Average Tree Height per House",
       title = "Average Tree Height per House Across Economic Groups")

avg_tree_height_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-2_average-tree-height-per-house-vs-economic-level.png"),
      plot = avg_tree_height_economic_plot,
      height = 6,
      width = 6)
```