# Results

```{r message = FALSE, warning = FALSE}
# Loading libraries and data
library(readxl)
library(janitor)
library(tidyverse)
library(broom)
library(purrr)
library(cowplot)
library(car)

GARDEN_FILE_PATH <- "data/raw/GARDENS_spreadsheet.xlsx"
garden_data <- read_excel(GARDEN_FILE_PATH, sheet='Everything_statistics') |>
  clean_names() # clean column names

BIRD_FILE_PATH <- "data/raw/BIRDS_spreadsheet2.xlsx"
bird_data <- read_excel(BIRD_FILE_PATH, sheet="Long_clean") |>
  clean_names()

# File path to save images of the graphs to
IMAGE_FILE_PATH <-"images/"
```

## Chapter 1

### Canopy cover in backyards vs. Economic Level

```{r}
canopy_ecomomic_lr <- lm(canopy_cover_garden_maybepercentage ~ economic_level, garden_data)
results_canopy_ecomomic_lr <- summary(canopy_ecomomic_lr)
results_canopy_ecomomic_lr
```

```{r}
canopy_ecomomic_lr_pvalue <- results_canopy_ecomomic_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
canopy_ecomomic_lr_r2 <- results_canopy_ecomomic_lr$r.squared |>
  round(4)

cat("p-value:", canopy_ecomomic_lr_pvalue, "\n")
cat("R-squared:", canopy_ecomomic_lr_r2)
```

```{r message = FALSE}
canopy_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = canopy_cover_garden_maybepercentage)) +
  geom_smooth(method = "lm") +
  geom_point(data = garden_data[-17, ]) +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Canopy Cover Garden (%)",
       title = "Backyard Canopy Cover vs. Economic Level")

canopy_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-14_canopy-cover-vs-economic-level.png"),
       plot = canopy_ecomomic_plot)
```

### Tree total biomass per backyard vs. Economic Level

```{r}
tree_biomass_ecomomic_lr <- lm(biomass_total_kg ~ economic_level, garden_data)
results_tree_biomass_ecomomic_lr <- summary(tree_biomass_ecomomic_lr)
results_tree_biomass_ecomomic_lr
```

```{r}
tree_biomass_ecomomic_lr_pvalue <- results_tree_biomass_ecomomic_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
tree_biomass_ecomomic_lr_r2 <- results_tree_biomass_ecomomic_lr$r.squared |>
  round(4)

cat("p-value:", tree_biomass_ecomomic_lr_pvalue, "\n")
cat("R-squared:", tree_biomass_ecomomic_lr_r2)
```

```{r, message = FALSE}
garden_biomass_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level, y = biomass_total_kg)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Garden Biomass (kg)",
       title = "Garden Biomass vs. Economic Level")

garden_biomass_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-15_garden-biomass-vs-economic-level.png"),
       plot = garden_biomass_ecomomic_plot)
```

### Total vegetation species richness per backyard vs Economic level

```{r}
veg_richness_ecomomic_lr <- lm(overall_garden_diversity ~ economic_level, garden_data)
results_veg_richness_lr <- summary(veg_richness_ecomomic_lr)
results_veg_richness_lr
```

```{r}
veg_richness_ecomomic_lr_pvalue <- results_veg_richness_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
veg_richness_lr_r2 <- results_veg_richness_lr$r.squared |>
  round(4)

cat("p-value:", veg_richness_ecomomic_lr_pvalue, "\n")
cat("R-squared:", veg_richness_lr_r2)
```

```{r, message = FALSE}
veg_richness_plot <-
  ggplot(garden_data, aes(x = economic_level, y = overall_garden_diversity)) +
  geom_smooth(method = "lm") +
  geom_point(data = garden_data[-1, ]) +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Total Backyard Plant Species Richness",
       title = "Total Backyard Plant Species Richness vs. Economic Level")

veg_richness_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-16_veg-richness-vs-economic-level.png"),
       plot = veg_richness_plot)
```

### Tree diversity in tems of percentage area per backyeard vs. Economic Level

```{r}
shannon_ecomomic_lr <- lm(shannon_index_per_garden_withpercentcover ~ economic_level, garden_data)
results_shannon_ecomomic_lr <- summary(shannon_ecomomic_lr)
results_shannon_ecomomic_lr
```

```{r}
shannon_ecomomic_lr_pvalue <- results_shannon_ecomomic_lr$coefficients["economic_level", "Pr(>|t|)"] |>
  round(4)
shannon_biomass_ecomomic_lr_r2 <- results_shannon_ecomomic_lr$r.squared |>
  round(4)

cat("p-value:", shannon_ecomomic_lr_pvalue, "\n")
cat("R-squared:", shannon_biomass_ecomomic_lr_r2)
```

```{r, message = FALSE}
shannon_ecomomic_plot <-
  ggplot(garden_data, aes(x = economic_level,
                          y = shannon_index_per_garden_withpercentcover)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Economic Level",
       y = "Shannon Diversity Index",
       title = "Shannon Diversity in Terms of Canopy Percent Cover vs. Economic Level")

shannon_ecomomic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-18_shannon-diversity-vs-economic-level.png"),
       plot = shannon_ecomomic_plot)
```

### ANOVA on how economic level influences tree species richness
```{r}
# add economic group level names to data
garden_data <- garden_data |>
  mutate(economic_category = as.factor(str_sub(house, 1, 1)),
         economic_category = as.factor(
           case_when(
             economic_category == "P" ~ "Low",
             economic_category == "M" ~ "Medium",
             economic_category == "R" ~ "High"
             )
           ),
         # specify order of economic group levels
         economic_category = fct_relevel(economic_category, c("Low", "Medium", "High")))
```

ANOVA:
```{r}
tree_richness_economic_anova <- aov(tree_species_richness ~ economic_category, garden_data)
results_tree_richness_economic_anova <- summary(tree_richness_economic_anova)
results_tree_richness_economic_anova
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(tree_richness_economic_anova$residuals)
```

```{r}
leveneTest(tree_species_richness ~ economic_category, garden_data)
```
Neither assumption is violated (p-values > 0.05).

**Summary of ANOVA results:**
```{r}
tree_richness_economic_anova_df <-
  results_tree_richness_economic_anova[[1]][[1, "Df"]] |>
  round(4)

tree_richness_economic_anova_df_resid <-
  results_tree_richness_economic_anova[[1]][[2, "Df"]] |>
  round(4)

tree_richness_economic_anova_fvalue <-
  results_tree_richness_economic_anova[[1]][[1, "F value"]] |>
  round(4)

tree_richness_economic_anova_pvalue <- 
  results_tree_richness_economic_anova[[1]][[1, "Pr(>F)"]] |>
  round(4)


cat("df (degrees of freedom):", tree_richness_economic_anova_df, "\n")
cat("Residual:", tree_richness_economic_anova_df_resid, "\n")
cat("F:", tree_richness_economic_anova_fvalue, "\n")
cat("p-value:", tree_richness_economic_anova_pvalue)
```

Plot:
```{r fig.height = 5, fig.width = 5}
tree_richness_economic_plot <-
  ggplot(garden_data, aes(x = economic_category, y = tree_species_richness)) +
  geom_boxplot() +
  theme_minimal() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  labs(x = "Economic Group",
       y = "Tree Species Richness",
       title = "Tree Species Richness Across Economic Groups")

tree_richness_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "1-19_tree-richness-vs-economic-level.png"),
       plot = tree_richness_economic_plot,
       height = 5,
       width = 5)
```

## Chapter 2

### Bird species richness in each backyard vs % canopy cover in back yard
removed outlier in analysis
```{r}
# (omit outlier)
bird_canopy_lr <- lm(bird_diversity ~ canopy_cover_garden_maybepercentage, garden_data[-13, ])
results_bird_canopy_lr <- summary(bird_canopy_lr)
results_bird_canopy_lr
```

```{r}
bird_canopy_lr_pvalue <-
  results_bird_canopy_lr$coefficients["canopy_cover_garden_maybepercentage", "Pr(>|t|)"] |>
  round(4)
bird_canopy_lr_r2 <- results_bird_canopy_lr$r.squared |>
  round(4)

cat("p-value:", bird_canopy_lr_pvalue, "\n")
cat("R-squared:", bird_canopy_lr_r2)
```

```{r, message = FALSE}
bird_canopy_plot <-
  ggplot(garden_data[-13, ], aes(x = canopy_cover_garden_maybepercentage,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Canopy Cover Backyard (%)",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Canopy Cover")

bird_canopy_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-21_bird_vs_canopy.png"),
       plot = bird_canopy_plot)
```

### Bird species richness vs vegetation species richness
kept outlier in analysis
```{r}
# (omit outlier)
bird_veg_lr <- lm(bird_diversity ~ overall_garden_diversity, garden_data)
results_bird_veg_lr <- summary(bird_veg_lr)
results_bird_veg_lr
```

```{r}
bird_veg_lr_pvalue <-
  results_bird_veg_lr$coefficients["overall_garden_diversity", "Pr(>|t|)"] |>
  round(4)
bird_veg_lr_r2 <- results_bird_veg_lr$r.squared

cat("p-value:", bird_veg_lr_pvalue, "\n")
cat("R-squared:", bird_veg_lr_r2)
```

```{r, message = FALSE}
bird_veg_plot <-
  ggplot(garden_data[-13, ], aes(x = overall_garden_diversity,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Total Plant Species Richness in Backyard",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Backyard Plant Diversity")

bird_veg_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-22_bird_vs_veg_richness.png"),
       plot = bird_veg_plot)
```

### Bird species richness vs native plant species as a ratio of all species in backyard
didn't include outlier in analysis
```{r message = FALSE}
# add native percentage to garden data
native_percent <- read_excel(GARDEN_FILE_PATH, sheet='Native_vs_not_GRAPHING') |>
  select(House, Native_percent_total) |>
  clean_names() # clean column names

garden_data <- garden_data |>
  left_join(native_percent, by = join_by(house))
```
```{r}
bird_native_lr <- lm(bird_diversity ~ native_percent_total, garden_data[-13, ])
results_bird_native_lr <- summary(bird_native_lr)
results_bird_native_lr
```

```{r}
bird_native_lr_pvalue <-
  results_bird_native_lr$coefficients["native_percent_total", "Pr(>|t|)"] |>
  round(4)
bird_native_lr_r2 <- results_bird_native_lr$r.squared

cat("p-value:", bird_native_lr_pvalue, "\n")
cat("R-squared:", bird_native_lr_r2)
```
Limit y-axis to 0-30?

```{r, message = FALSE}
bird_native_plot <-
  ggplot(garden_data[-13, ], aes(x = native_percent_total,
                          y = bird_diversity)) +
  geom_smooth(method = "lm") +
  geom_point() +
  theme_minimal() +
  labs(x = "Native % Plants Total",
       y = "Bird Species Richness",
       title = "Bird Species Richness vs. Native Plant Proportion")

bird_native_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-23_bird_vs_native.png"),
       plot = bird_native_plot)
```

### Accumulation curve accross houses
```{r}
set.seed(123) # set seed for reproducible permutations

# presence/absence matrix where rows = houses and columns = species
presence_matrix <- bird_data |>
  count(house, species) |>
  pivot_wider(names_from = species, values_from = n, values_fill = 0) |>
  column_to_rownames("house") |>
  mutate(across(everything(), ~ as.integer(. > 0)))

# species accumulation curve
n_permutations <- 100
houses <- rownames(presence_matrix)
n_houses <- length(houses)
n_species <- ncol(presence_matrix)

curves <- matrix(0, nrow = n_permutations, ncol = n_houses)

for (i in seq_len(n_permutations)) {
  perm <- sample(houses)
  seen <- rep(FALSE, n_species)

  for (j in seq_along(perm)) {
    seen <- seen | as.logical(presence_matrix[perm[j], ])
    curves[i, j] <- sum(seen)
  }
}

# summarise across permutations
results <- tibble(
  x          = seq_len(n_houses),
  mean_curve = colMeans(curves),
  lower      = apply(curves, 2, quantile, probs = 0.025),
  upper      = apply(curves, 2, quantile, probs = 0.975)
)

# plot
accumulation_plot <- ggplot(results, aes(x = x)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey60", alpha = 0.4) +
  geom_line(aes(y = mean_curve), color = "#3366FF", linewidth = 1) +
  labs(
    title = "Species Accumulation Curve Across Houses",
    x     = "Number of Houses Sampled",
    y     = "Cumulative Species Detected"
  ) +
  theme_minimal()

accumulation_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-24_accumulation_plot.png"),
       plot = accumulation_plot)
```

### Panel comparing vegetation and birds variables vs economic level and the same over area of backyard
```{r fig.height = 12, fig.width = 6}
garden_data <- garden_data |>
  mutate(biomass_per_m2 = biomass_total_kg/garden_size_m_2,
         plants_per_m2 = overall_garden_diversity/garden_size_m_2,
         trees_per_m2 = tree_species_richness/garden_size_m_2,
         canopy_per_m2 = canopy_cover_garden_maybepercentage/garden_size_m_2,
         birds_per_m2 = bird_diversity/garden_size_m_2,
         )

y_vars <- c("biomass_total_kg", "biomass_per_m2",
            "overall_garden_diversity", "plants_per_m2",
            "tree_species_richness", "trees_per_m2",
            "canopy_cover_garden_maybepercentage", "canopy_per_m2",
            "bird_diversity", "birds_per_m2")
titles <- c("Biomass", "Biomass per area",
            "Plant richness", "Plant richness per area",
            "Tree richness", "Tree richness per area",
            "Canopy cover", "Canopy cover per area",
            "Bird richness", "Bird richness per area")

plots <- map2(y_vars, titles, \(var, title) {
  # compute p-value for the slope of economic_level
  p <- tidy(lm(garden_data[[var]] ~ garden_data$economic_level))$p.value[2]
  p_label <- paste0("p = ", signif(p, 3))
  
  ggplot(garden_data, aes(x = economic_level, y = .data[[var]])) +
    geom_smooth(method = "lm") +
    geom_point() +
    theme_minimal() +
    annotate("text", x = -Inf, y = Inf, label = p_label,
             hjust = -0.1, vjust = 1.5, size = 4) +
    labs(y = NULL, x = NULL, title = title)
})

panel_plot <- plot_grid(plotlist = plots, ncol = 2)
panel_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "2-25_panel-plot.png"),
       plot = panel_plot,
       height = 12,
       width = 6)
```
## Other

### ANOVA tree height per economic level
```{r}
tree_height <- read_excel(GARDEN_FILE_PATH, sheet='Tree_per_garden') |>
  select(ID, `Ht (m)`) |>
  clean_names() |>
  left_join(garden_data[, c("house", "economic_category")], by = c("id" = "house")) |>
  filter(id != "P22") # remove outlier
```

ANOVA:
```{r}
tree_height_economic_anova <- aov(ht_m ~ economic_category, tree_height)
results_tree_height_economic_anova <- summary(tree_height_economic_anova)
results_tree_height_economic_anova
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(tree_height_economic_anova$residuals)
```

```{r}
leveneTest(ht_m ~ economic_category, tree_height)
```
The assumption of normally distributed residuals is violated (p-value < 0.05) while the assumption of variance is not (p-value > 0.05). Therefore, a Kruskal-Wallis test is more appropriate to compare the **median** of each group instead of the **mean**. The medians better reflects the central tendency in the case of outliers and skewedness. Additionally, the Kruskal-Wallis test does not make assumptions about the data distribution or variablility.

Kruskal-Wallis Test:
```{r}
kruskal.test(ht_m ~ economic_category, data = tree_height)
```
There is a statistically significant difference in median tree height across economic categories (p = 0.022 < 0.05).

```{r fig.height = 6, fig.width = 5}
tree_height_economic_plot <-
  ggplot(tree_height, aes(x = economic_category, y = ht_m)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Group",
       y = "Tree Height",
       title = "Tree Height Across Economic Groups")

tree_height_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-1_tree-height-vs-economic-level.png"),
      plot = tree_height_economic_plot,
      height = 6,
      width = 6)
```
  
### ANOVA average tree height per house per economic level
```{r}
average_tree_height_per_house <- tree_height |>
  group_by(id, economic_category) |>
  summarize(average_tree_height = mean(ht_m))

avg_tree_height_economic_anova <-
  aov(average_tree_height ~ economic_category, average_tree_height_per_house)

summary(avg_tree_height_economic_anova)
```

**Check assumptions (1) normality of residuals using a Shapiro-Wilk normality test and (2) equal variance of each group using a Levene's test.**
```{r}
shapiro.test(avg_tree_height_economic_anova$residuals)
```

```{r}
leveneTest(average_tree_height ~ economic_category, data = average_tree_height_per_house)
```
Again, the assumption of normally distributed residuals is violated (p-value < 0.05) while the assumption of variance is not (p-value > 0.05). Therefore, a Kruskal-Wallis test again to compare the **median** of each group instead of the **mean**.

Kruskal-Wallis Test:
```{r}
kruskal.test(average_tree_height ~ economic_category, data = average_tree_height_per_house)
```
The median of average tree height per house did not differ significantly across economic categories (p-value > 0.064 > 0.05).


```{r fig.height = 6, fig.width = 5}
avg_tree_height_economic_plot <-
  ggplot(average_tree_height_per_house, aes(x = economic_category, y = average_tree_height)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  theme_minimal() +
  labs(x = "Economic Group",
       y = "Average Tree Height per House",
       title = "Average Tree Height per House Across Economic Groups")

avg_tree_height_economic_plot

ggsave(filename = paste0(IMAGE_FILE_PATH, "other-2_average-tree-height-per-house-vs-economic-level.png"),
      plot = avg_tree_height_economic_plot,
      height = 6,
      width = 6)
```